---
title: "Assignment 1"
output: html_notebook
---

Question 1 (40 points)

According to the National Association of Colleges and Employers, the average salary for a new college graduate in 2019 was $45,000. A small academic program wants to know if the average salary of its graduates is below the national mean value in 2019. If it is below the national mean value, they want to take measures to increase the starting salary of their graduates. The starting salary of 36 recent graduates of this program is stored in an R vector called “college_salary”. Copy the following text, paste it in an R studio code chunk, and run the code chunk. Afterward, you will have the vector available in your R project.

```{r}
college_salary= c (45600, 51400, 35600, 42892, 48180, 51000, 39900, 40000, 48100, 42033 ,41072, 42898, 42010, 51200, 43260, 42901, 40296, 44350, 43200, 47512, 45590, 50050, 41110, 42194, 44643, 43600, 40600, 50961, 39254, 42400, 52000, 46100, 50533, 42026, 41236, 44239)
```

a) (25 points) Conduct the corresponding hypothesis test. 

```{r}
t.test(college_salary, alternative= c("less"), mu= 45000)

```
Report the following as part of your answer:

The hypotheses (Ho and Ha).

Null Hypothesis (H₀): The mean salary of the program's graduates is equal to or greater than the national average salary.
H0: μ = 45000

Alternative Hypothesis (Hₐ): The mean salary of the program's graduates is less than the national average salary.
Ha: μ < 45000

---
The P value.

P Value = 0.2115
---
Your conclusion.

Given that the p-value is greater than alpha of 0.05, we do not have sufficient evidence to reject the null hypothesis.

Conclusion: There is not enough statistical evidence to conclude that the average salary of this program's graduates is below the national mean of $45,000. Therefore, we cannot say that the program's average salary is significantly lower than the national average.
-------------------------------------------------------------------------------------------

b) (10 points) Show the computations needed to get the value of the test statistic that you obtained in the R output from part a. In other words, you need to show how that value was obtained.

``` {r}
#sample mean
sampleMean = mean(college_salary)

#population mean given as $45,000
popMean = 45000

#sample standard Deviation
sampleSD = sd(college_salary)

# of samples
sample_count = length(college_salary)

#t-value
tvalue = ((sampleMean - popMean)/(sampleSD/(sqrt(sample_count))))
print(tvalue)

```


-------------------------------------------------------------------------------------------
c) (5 points) What type of statistical error could you potentially be making after your decision in part a? Justify.

Since the conclusion was not to reject the null hypothesis, the possible error is a Type II error, which implies that we might be incorrectly concluding there is no significant difference when, in reality, the program's average salary could be below the national mean.
-------------------------------------------------------------------------------------------

Question 2 (60 points) 
Researchers at a Zoo wish to test the efficacy of a new program intended to reduce the length of labor in female pumas (i.e., mountain lions). From historical records, researchers know that it takes an average of 60 minutes for a female puma to give birth to a litter. The new program was tested with 25 puma mothers that gave birth. The data of the length of labor (in minutes) for these 25 puma mothers is shown below.

```{r}
puma_labor= c(56, 51, 62, 58, 59, 59, 56, 56, 54, 47, 51, 64, 48, 59, 59, 53, 61, 55, 57, 63, 55, 56, 62, 57, 55)
```


---
a) (10 points) How many standard deviations from the sample average is the most extreme value in the sample? Show your work and make sure you provide a written answer to the question.

```{r}
#Sample Mean
q2SampleMean = mean(puma_labor)

#find extreme value of sample
extremeValue = puma_labor[which.max(abs(puma_labor - q2SampleMean))]
print(extremeValue)

#find the Standard Deviation of the sample
q2sampleSD = sd(puma_labor)

#calculate sd from sample average 
numSDofX = (extremeValue - q2SampleMean) / q2sampleSD
print(numSDofX)

```


The most extreme value in the sample is approximately 2.18 standard deviations away from the sample mean. This means that this value deviates from the average by a bit more than twice the standard deviation of the sample.
---

b) (35 points) Test whether the mean labor time attained with this new program is less than the historical mean. 
```{r}
q2TestResult = t.test(puma_labor, alternative= c("less"), mu= 60)
q2TestResult
```

Report the following as part of your answer:
---
The hypotheses (Ho and Ha).

Null Hypothesis (H₀): The mean labor time with the new program is equal to the historical mean.
H0: μ = 60

Alternative Hypothesis (Hₐ): The mean labor time with the new program is less than the historical mean.
Ha: μ < 60

---
The P value.
```{r}
pValue = q2TestResult$p.value
pValue
```


---
Your conclusion.

Since the p-value (0.00027) is much smaller than the common significance level of 0.05, we reject the null hypothesis. This means there is significant evidence to suggest that the mean labor time for pumas under the new program is less than the historical mean of 60 minutes. Therefore, the new program appears to be effective in reducing the length of labor in female pumas.
---

c) (10 points) (This is the challenge question)

Note: Challenge questions are those I want you to answer without getting any help from me (i.e., you cannot ask me how to answer them). It is only 10 points; thus, do not worry. Try your best to answer it.

What type of statistical error could you potentially be making after your decision in part c? Is the probability that you are making this type of error small or large? Justify the answers to both of these questions.

Since the null hypothesis is being rejected, the potential for a Type I statistical error exists. However, since the p-value of 0.000269 is much smaller than the alpha of 0.05, the probability of making this error is small and extremely unlikely.
---
d) (5 points) Would a sample average of 61 minutes for the labor time give you more or less evidence against Ho? Justify your answer.

A sample mean of 61 would give less evidence against Ho, however not in a very meaningful manner as the number of samples and their standard deviation is what drives the results of the test.
